---
title: "Diamond price prediction"
author: "Kerim Kiliç"
subtitle: Supervised Machined Learning
output:
  html_document:
    df_print: paged
---

# Libraries

The following three libraries are used in this R notebook.

```{r setup, message=FALSE}
library(tidymodels)
library(GGally)
library(gridExtra)
```

# Dataset descriptive analytics

Glimpse into the variables of the *diamonds* dataset.

```{r}
diamonds %>% glimpse()
```

Take a smaller sample from dataset to reduce time to train the model.
```{r}
diamonds_sample <- diamonds %>% slice_sample(prop = 0.2)
```

ggpairs plot of all the variables and their correlation in the dataset. Set *eval* to *TRUE* to create this plot.

```{r, out.width='100%', fig.height=7, eval=FALSE}
ggpairs(diamonds_sample,
        lower = list(combo = wrap("facethist", bins = 10)))
```

ggpairs plot of all the highly correlated variables in the dataset. Set *eval* to *TRUE* to create this plot.

```{r, out.width='100%', fig.height=7, eval=FALSE}
ggpairs(diamonds_sample %>% select(carat, price:z),
        lower = list(combo = wrap("facethist", bins = 10)))
```

Histograms with the distribution of the target variable *price*.

```{r}
plot3 <- ggplot(diamonds_sample, aes(price)) +
  geom_histogram(bins = 20) +
  theme_minimal()
plot4 <- ggplot(diamonds_sample, aes(price)) +
  geom_histogram(bins = 20) +
  scale_x_log10() +
  theme_minimal()
grid.arrange(plot3,plot4,ncol=2)
```

# Variable transformation

Create a variable with the logarithm of the target variable.

```{r}
diamonds <- diamonds %>%
  mutate(log_price = log10(price))
```


# Data split

Split between the train set and the test set.

```{r}
set.seed(2022)
diamonds_split <- initial_split(diamonds, prop = 0.8, strata = log_price)
```

# Recipes

Create recipe.

```{r}
initial_recipe <- training(diamonds_split) %>%
  recipe(log_price ~ .) %>%
  step_rm(price) %>%
  step_poly(carat, degree = 2) %>%
  step_dummy(all_nominal(), one_hot = TRUE) %>%
  step_corr(all_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())

initial_recipe
```

Glimpse into recipe.

```{r}
initial_recipe %>% prep() %>% juice() %>% glimpse()
```

# Creating models

## Regression based models

Show engines of regression based models:

```{r}
show_engines("linear_reg")
```

linear regression: lm

```{r}
lm <- linear_reg(mode = "regression") %>%
  set_engine("lm")
```

# Testing the models

## Metric set and folds

```{r}
metric_diamonds <- metric_set(rsq, mae, rmse)
```

```{r}
folds <- vfold_cv(training(diamonds_split), v = 5, repeats = 2)
```

## Cross validating our linear model

```{r}
wf_lm <- workflow() %>%
  add_recipe(initial_recipe) %>%
  add_model(lm)
```

```{r, message=FALSE}
fit_resamples(wf_lm, 
              folds,
              metrics = metric_diamonds) %>%
  collect_metrics()
```

