---
title: "Diamond price prediction"
subtitle: "Supervised Machined Learning"
author: "Kerim Kiliç"
output:
  html_document:
    df_print: paged
---

The following three libraries are used in this R notebook.

```{r setup, message=FALSE}
library(tidymodels)
library(GGally)
library(gridExtra)
```

Glimpse into the variables of the *diamonds* dataset.

```{r}
diamonds %>% glimpse()
```

Take a smaller sample from dataset to reduce time to train the model.
```{r}
diamonds_sample <- diamonds %>% slice_sample(prop = 0.2)
```

ggpairs plot of all the variables and their correlation in the dataset. Set *eval* to *TRUE* to create this plot.

```{r, out.width='100%', fig.height=7, eval=FALSE}
ggpairs(diamonds_sample,
        lower = list(combo = wrap("facethist", bins = 10)))
```

ggpairs plot of all the highly correlated variables in the dataset. Set *eval* to *TRUE* to create this plot.

```{r, out.width='100%', fig.height=7, eval=FALSE}
ggpairs(diamonds_sample %>% select(carat, price:z),
        lower = list(combo = wrap("facethist", bins = 10)))
```

Histograms with the distribution of the target variable *price*.

```{r}
plot3 <- ggplot(diamonds_sample, aes(price)) +
  geom_histogram(bins = 20) +
  theme_minimal()
plot4 <- ggplot(diamonds_sample, aes(price)) +
  geom_histogram(bins = 20) +
  scale_x_log10() +
  theme_minimal()
grid.arrange(plot3,plot4,ncol=2)
```